Tidycensus Markdown
Trevor Kapuvari
9/4/2023

# Question 1 Lab Markdown 
```{r setup, include=FALSE, cache = TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(kableExtra)
library(rmarkdown)
```

```{r setup_packages1, warning = FALSE, eval = FALSE}
install.packages('tidyverse')
```

```{r setup_packages2, warning = FALSE, message = FALSE}
library(tidyverse)
library(tidycensus)
library(sf)
```

```{r load_key, warning = FALSE, eval = FALSE}
census_api_key("2ad9e737f3d9062836cb46bb568be5467f86d3db", overwrite = TRUE)


{r load_variables, cache = TRUE}

acs_variable_list.2020 <- load_variables(2020, cache = TRUE)

acs_variable_list.2016 <- load_variables(2016, cache = TRUE)
```

```{r acs_vars, include=FALSE}
acs_vars <- c("B01001_001E", # ACS total Pop estimate
              "B25002_001E", # Estimate of total housing units
              "B25002_003E", # Number of vacant housing units
              "B19013_001E", # Median HH Income ($)
              "B02001_002E", # People describing themselves as "white alone"
              "B06009_006E") # Total graduate or professional degree
```

```{r get_acs_2020, message=FALSE, warning=FALSE, cache=TRUE, include=FALSE}
acsTractsPHL.2020 <- get_acs(geography = "tract",
                             year = 2020, 
                             variables = acs_vars, 
                             geometry = FALSE, 
                             state = "PA", 
                             county = "Philadelphia", 
                             output = "wide") 
```

```{r do_some_dplyr, cache=FALSE, include=FALSE}
acsTractsPHL.2020 <- acsTractsPHL.2020 %>%
  dplyr::select (GEOID, NAME, all_of(acs_vars))

acsTractsPHL.2020 <- acsTractsPHL.2020 %>%
  rename (total_pop.2020 = B01001_001E,
          total_HU.2020 = B25002_001E,
          total_vacant.2020 = B25002_003E,
          med_HH_Income.2020 = B19013_001E,
          total_White.2020 = B02001_002E,
          total_GradDeg.2020 = B06009_006E)

acsTractsPHL.2020 <- acsTractsPHL.2020 %>%
  mutate(vacancyPct.2020 = total_vacant.2020/total_HU.2020,
         pctWhite.2020   = total_White.2020/total_pop.2020)
```

```{r get_acs_2016, message=FALSE, cache=TRUE, include=FALSE}
acsTractsPHL.2016 <- get_acs(geography = "tract",
                             year = 2016, 
                             variables = acs_vars,
                             geometry = FALSE,
                             state = "PA", 
                             county = "Philadelphia",
                             output = "wide") %>%
  dplyr::select (GEOID, NAME, all_of(acs_vars)) %>% 
  rename (total_pop.2016 = B01001_001E,
          total_HU.2016 = B25002_001E,
          total_vacant.2016 = B25002_003E,
          med_HH_Income.2016 = B19013_001E,
          total_White.2016 = B02001_002E,
          total_GradDeg.2016 = B06009_006E) %>%
  mutate(vacancyPct.2016 = total_vacant.2016/total_HU.2016,
         pctWhite.2016 = total_White.2016/total_pop.2016)
```

```{r left_join_tracts, cache=FALSE, include=FALSE}
allACS <- left_join(acsTractsPHL.2016, acsTractsPHL.2020,
                    by= c("GEOID"))
```

```{r do_mutates, cache=FALSE, include=FALSE}
allACS <- allACS %>%
  mutate(change_med_HH_Income = med_HH_Income.2020 - (med_HH_Income.2016 * 1.08), 
         change_Grad_Degree_Pct = (total_GradDeg.2020/total_pop.2020)-(total_GradDeg.2016/total_pop.2016))

```

```{r base_r_summaries_NA_example, cache=FALSE, include=FALSE}
mean(allACS$change_med_HH_Income)
```

```{r base_r_summaries_NA_rm, cache=FALSE, include=FALSE}
mean(allACS$change_med_HH_Income, na.rm = TRUE)

median(allACS$change_med_HH_Income, na.rm = TRUE)

```

```{r basic_histogram, cache=FALSE, include=FALSE}
hist(allACS$change_med_HH_Income)
```

```{r ggplot_histogram_simple, message=FALSE, warning=FALSE, cache=TRUE, include=FALSE}
ggplot(allACS) + 
  geom_histogram(aes(change_med_HH_Income))
  
```

```{r ggplot_histogram, warning=FALSE, cache=FALSE, include=FALSE}
ggplot(allACS)+
  geom_histogram(aes(change_med_HH_Income), binwidth = 5000)+
  labs(
    title = "Change in Philadelphia HH median income by tract, 2016-2020",
    caption = "Data: US Census Bureau, ACS 5-year estimates",
    x="Change in Med HH Income (2020 dollars)", 
       y="Number of tracts")
  
```

```{r summaryTable, cache=FALSE, include=FALSE}
summaryTable <- allACS %>%
  summarize(mean_change_HH_Income = mean(change_med_HH_Income, na.rm = TRUE),
            med_change_HH_Income = median(change_med_HH_Income, na.rm = TRUE))
```

```{r myTracts, cache=FALSE, include=FALSE}

myTracts <- c("42101023500", 
              "42101023600", 
              "42101023700", 
              "42101025300", 
              "42101025400",
              "42101025500", 
              "42101025600", 
              "42101038800")

allACS <- allACS %>%
  mutate(mtAiry = ifelse(GEOID %in% myTracts, "MT AIRY", "REST OF PHILADELPHIA"))
```

```{r summary_table_2, message=FALSE, warning=FALSE, cache=TRUE, include=FALSE}
summaryTable2 <- allACS %>%
  group_by(mtAiry) %>%
  summarize(mean_change_HH_Income = mean(change_med_HH_Income, na.rm = TRUE),
            med_change_HH_Income = median(change_med_HH_Income, na.rm = TRUE))
```

```{r ggplot_histogram_2, warning=FALSE, cache=FALSE, include=FALSE}
ggplot(allACS)+
  geom_histogram(aes(change_med_HH_Income),
                 binwidth = 5000)+
  labs(
    title = "Change in Philadelphia HH median income by tract, 2016-2020",
    caption = "Data: US Census Bureau, ACS 5-year estimates",
    x="Change in Med HH Income (2020 dollars)", 
       y="Number of tracts")+
  facet_wrap(~mtAiry, scales = "free")
  
```

```{r ggplot_point1, warning=FALSE, cache=FALSE, include=FALSE}
ggplot(allACS)+
  geom_point(aes(x =med_HH_Income.2016 * 1.08, 
                 y = med_HH_Income.2020,
                 color = mtAiry))+
  geom_abline(intercept = 0, slope = 1)+
  labs(
    title = "2020 Median HH Income as a Function of 2016 Median HH Income",
    subtitle = "All figures in 2020 dollars",
    caption = "Data: US Census Bureau, ACS 5-year estimates",
    x="Med HH Income 2016 ($)", 
    y="Med HH Income 2020 ($)")
  
```

```{r ggplot_point2, message=FALSE, warning=FALSE, cache=TRUE, include=FALSE}
ggplot(allACS)+
  geom_point(aes(x = 100* pctWhite.2020, 
                 y = med_HH_Income.2020,
                 color = mtAiry))+
  geom_smooth(aes(x = 100* pctWhite.2020, 
                  y = med_HH_Income.2020), 
              method = "lm", se = FALSE)+
  labs(
    title = "2020 Median HH Income as a Function of Pct White",
    subtitle = "All figures in 2020 dollars",
    caption = "Data: US Census Bureau, ACS 5-year estimates",
    x="Pct. Residents Identifying as 'White Only'", 
    y="Med HH Income 2020 ($)")
  
```

```{r spatial_tidycensus, message=FALSE, warning=FALSE, cache=TRUE, include=FALSE}
acsTractsPHL.2020.sf <- get_acs(geography = "tract",
                             year = 2020, 
                             variables = acs_vars, 
                             geometry = TRUE, 
                             state = "PA", 
                             county = "Philadelphia", 
                             output = "wide") %>% 
  dplyr::select (GEOID, NAME, all_of(acs_vars)) %>%
  rename (total_pop.2020 = B01001_001E,
          total_HU.2020 = B25002_001E,
          total_vacant.2020 = B25002_003E,
          med_HH_Income.2020 = B19013_001E,
          total_White.2020 = B02001_002E,
          total_GradDeg.2020 = B06009_006E) %>%
  mutate(vacancyPct.2020 = total_vacant.2020/total_HU.2020,
         pctWhite.2020 = total_White.2020/total_pop.2020) %>%
  mutate(mtAiry = ifelse(GEOID %in% myTracts, "MT AIRY", "REST OF PHILADELPHIA")) %>%
  st_as_sf(crs = 4326) # Turn shp into sf object and project as WGS84
```

```{r spatial_tidycensus_no_eval, message=FALSE, warning=FALSE, cache=TRUE, include=FALSE}
acsTractsPHL.2020.sf <- get_acs(geography = "tract",
                             year = 2020, 
                             variables = acs_vars, 
                             geometry = TRUE, 
                             state = "PA", 
                             county = "Philadelphia", 
                             output = "wide") %>% 
  dplyr::select (GEOID, NAME, all_of(acs_vars)) %>%
  rename (total_pop.2020 = B01001_001E,
          total_HU.2020 = B25002_001E,
          total_vacant.2020 = B25002_003E,
          med_HH_Income.2020 = B19013_001E,
          total_White.2020 = B02001_002E,
          total_GradDeg.2020 = B06009_006E) %>%
  mutate(vacancyPct.2020 = total_vacant.2020/total_HU.2020,
         pctWhite.2020 = total_White.2020/total_pop.2020) %>%
  mutate(mtAiry = ifelse(GEOID %in% myTracts, "MT AIRY", "REST OF PHILADELPHIA")) %>%
  st_as_sf(crs = 4326) # Turn shp into sf object and project as WGS84
```


```{r ggplot_geom_sf2, warning=FALSE, cache=FALSE, include=FALSE}
ggplot()+
  geom_sf(data = acsTractsPHL.2020.sf, aes(fill = total_GradDeg.2020),
          color = "transparent")+
  geom_sf(data = acsTractsPHL.2020.sf %>%
            filter(mtAiry == "MT AIRY") %>%
            st_union(),
          color = "white",
          fill = "transparent")+
  labs(
    title = "Graduate Degrees",
    subtitle = "",
    caption = "Data: US Census Bureau, ACS 5-year estimates")
  
```

# Part 2 Questions and Code 

```{r ggplot_point3, warning = FALSE, cache = TRUE}
# Question 2 Section A, plot points
allACS <- allACS %>%
  mutate(change_total_vacant = total_vacant.2020 - total_vacant.2016)
  # Negative numbers show a decrease in vacant housing units 

ggplot(allACS)+
    geom_point(aes(x = vacancyPct.2020 * 100, 
                   y = change_total_vacant,
                   color = mtAiry))+
    geom_abline(intercept = 0, slope = 1)+
    labs(
        title = "2020 Vacancy Changes in Philadelphia",
        subtitle = "Compared to 2016",
        caption = "Data: US Census Bureau, ACS 5-year estimates",
        x="Percentage of Total Housing", 
        y="Total Vacancy Change")

```
With negative numbers representing decreased vacancy over time, Mt Airy (highlighted in orange) has a scattered pattern that shows a contrast between different parts of the neighborhood. 

```{r ggplot_geom_sf, warning = FALSE, cache = TRUE}
# Question 2 Section B, ggplot map 
acsTractsPHL.2016 <- get_acs(geography = "tract",
                              year = 2016, 
                              variables = acs_vars,
                              geometry = TRUE,
                              state = "PA", 
                              county = "Philadelphia",
                              output = "wide") %>%
     dplyr::select (GEOID, NAME, all_of(acs_vars)) %>% 
     rename (total_pop.2016 = B01001_001E,
             total_HU.2016 = B25002_001E,
             total_vacant.2016 = B25002_003E,
             med_HH_Income.2016 = B19013_001E,
             total_White.2016 = B02001_002E,
             total_GradDeg.2016 = B06009_006E) %>%
     mutate(vacancyPct.2016 = total_vacant.2016/total_HU.2016,
            pctWhite.2016 = total_White.2016/total_pop.2016)

myTracts <- c("42101023500", 
               "42101023600", 
               "42101023700", 
               "42101025300", 
               "42101025400",
               "42101025500", 
               "42101025600", 
               "42101038800")
 
 acsTractsPHL.2016 <- acsTractsPHL.2016 %>%
     mutate(mtAiry = ifelse(GEOID %in% myTracts, "MT AIRY", "REST OF PHILADELPHIA"))
# The above code was needed to re-add geometry and the Mt Airy data to the Phil2016 Table, originally only on the Phil2020 Table.
ggplot()+
    geom_sf(data = acsTractsPHL.2016, aes(fill = vacancyPct.2016 * 100),
            color = "transparent")+
    geom_sf(data = acsTractsPHL.2016 %>%
                filter(mtAiry == "MT AIRY") %>%
                st_union(),
            color = "white",
            fill = "transparent")+
    labs(
        title = "Vacancy Percentage 2016",
        subtitle = "",
        caption = "Data: US Census Bureau, ACS 5-year estimates")

```
The map visualizes how Mt.Airy has areas of both higher and lower than average vacancy rates among its tracts.  
```{r summaryTable4, cache = TRUE}
# Question 3 Section c, Kable Table 
summaryTable4 <- allACS %>%
    group_by(mtAiry) %>%
    summarize(mean_total_vacant.2020 = mean(total_vacant.2020, na.rm = TRUE),
              med_total_vacant.2020 = median(total_vacant.2020, na.rm = TRUE))
summaryTable4 %>%
  kbl() %>%
  kable_classic_2(full_width = F)
```
The kable table compares Mt Airy's average amount of vacant homes per tract to the rest of Philadelphia, showing that this neighborhood has, on average, a lower vacancy rate than the rest of the city. 